---
title: "Single RDO without dependencies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Single RDO without dependencies}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  cache = FALSE,
  comment = "#>"
)
```

## Creating a Reproducible Data Object (RDO)

Let's create a single RDO without any dependencies.
Our ROD object will be encapsulating `mtcars` dataset under different
name `data_mtcars` and the reproducible R code. 

```{r}
# library(RDO)
data_mtcars <- RDO::RDO$new(name = "data_mtcars")

data_mtcars
```

Printing the object gives you the name of the object,
dependencies if there are any, and statuses.

Let's add the reproducible code. The code needs to be of `expression` class.
We need simply to wrap our code in `expression({ })`.

```{r}
data_mtcars$code <- expression({
  data_mtcars <- mtcars
})

data_mtcars
```

If we changing the code or the cached data the timestamp `last changed` 
is updated. 
Reading the cache changes the `last touched` timestamp, so we know 
when the RDO was last used. 

We can access the cached data via active binding `cache`.

```{r}
data_mtcars$cache
```

The cache is `NULL` as the code has never been run yet 
and the data haven't been cached yet. Let's do it for the first time...

## Running the RDO R code

```{r}
data_mtcars$run()
```

If we don't want to see messages printed out to the console we can turn it off
in RDO methods using `verbose` parameter (default is `TRUE`).

```{r}
data_mtcars$run(verbose = FALSE)

head(data_mtcars$cache)

data_mtcars
```

Now the data cache is available.
We can see how big is the size of the RDO cache.

## Implicit validation

After the code was evaluated and the result was cached, 
the status `is validated?` changed to `TRUE`. 
This is called _implicit validation_.
We can confirm it with `is_validated()` method.

```{r}
data_mtcars$is_validated()
```

## Chaining the RDO methods

Many of RDO methods return `self` object so we can chain them.

```{r}
data_mtcars$run()$is_validated()
```

## RDO getters

We can access the name of the RDO using `get_name()` method.

```{r}
data_mtcars$get_name()
```

The `get_status` method returns the list of different statuses.

```{r}
data_mtcars$get_status()
```

## Getting the code

To access the code as a text we can use `code` active binding.
It returns the RDO (self) but before that it prints out to the console
the RDO code. 

```{r}
data_mtcars$code
```


The `print_code()` method can give you similar effect.

```{r}
data_mtcars$print_code()
```

If you want to get your code as R expressions you can use `get_code()` method.
The returned expression can be than evaluated manually. 
You can also use `run_code(cache = FALSE)` method for that purpose. 
The code will be evaluated but result will not be cached.

```{r}
data_mtcars$get_code()

eval_result <- eval(data_mtcars$get_code(), envir = new.env())
head(eval_result)

head(data_mtcars$run(cache = FALSE))
```

## Locking an RDO

You can reset your cache to by setting its value to `NULL`, 
but you cannot do it if the RDO is locked.

```{r error=TRUE}
data_mtcars$cache <- NULL

data_mtcars$run()$lock()$is_locked()

data_mtcars$cache <- NULL

data_mtcars$unlock()
```

## Explicit validation

Running the code and caching the result _implicitely validating_ the RDO,
but there is also an _explicit validation_. 
Explicit Validation is done using `validate()` method. 
The method is evaluating the reproducible code and the result is compare
to the cached data. If they are the same, the RDO is validated.

```{r}
data_mtcars$cache <- iris

data_mtcars$validate()$is_validated()

data_mtcars$cache <- mtcars

data_mtcars$validate()$is_validated()
```
