---
title: "RDO with dependencies"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RDO with dependencies}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Creating an RDO with dependencies

Let's start by creading a simple RDO that will be a dependency for another RDO.

```{r}
data_mtcars <- RDO::RDO$new(name = "data_mtcars")

data_mtcars$code <- expression({
  data_mtcars <- mtcars
})
```

Now let's create a new RDO with R code that needs the `data_mtcars` object
to be evaluated first. 

```{r}
mtcars_half_top <- RDO::RDO$new(name = "mtcars_half_top")

mtcars_half_top$code <- expression({
  mtcars_half_top <- head(data_mtcars, 16)
})
```

The code needs `data_mtcars` as dependency but no dependency is defined yet.

```{r}
mtcars_half_top$has_dependencies()
```

Running the RDO code will give an error as the dependecy is missing.

```{r, error=TRUE}
mtcars_half_top$run()
```

We can add add the missing dependency to the RDO.

```{r, error=TRUE}
mtcars_half_top$add_dependencies(dependencies = data_mtcars)

mtcars_half_top$run()
```

Running the code still gives an error. 
This time the dependency is present but is not validated.
Let's run the dependency code first.

```{r}
data_mtcars$run()
mtcars_half_top$run()
NROW(mtcars_half_top$cache)
mtcars_half_top$code
```

We can see that our `mtcars_half_top` RDO has `data_mtcars` RDO as dependency.
The object is valdiated, and the cache contains top 16 rows of `mtcars` dataset.

## Example RDO definitions

Let's define more RDOs with respective R code. 
We can use `dependencies` argument of the constructor 
function to set dependencies in one step.
 
```{r}
mtcars_half_bottom <-
  RDO::RDO$new(name = "mtcars_half_bottom",
               dependencies = list(data_mtcars))

mtcars_half_bottom$code <- expression({
  mtcars_half_bottom <- tail(data_mtcars, 16)
})

mtcars_whole <-
  RDO::RDO$new(name = "mtcars_whole",
               dependencies = list(mtcars_half_top,
                                   mtcars_half_bottom))

mtcars_whole$code <- expression({
  mtcars_whole <- rbind(mtcars_half_top,
                        mtcars_half_bottom)
})

data_iris <-
  RDO::RDO$new(name = "data_iris")

data_iris$code <- expression({
  data_iris <- iris
})

iris_selected_rows <-
  RDO::RDO$new(name = "iris_selected_rows",
               dependencies = list(data_iris))

iris_selected_rows$code <- expression({
  set.seed(1234)
  sampled_rows <- c(rep(TRUE, 32), rep(FALSE, NROW(data_iris) - 32))
  iris_selected_rows <- data_iris[sample(sampled_rows, NROW(data_iris)), ]
})

iris_selected_columns <-
  RDO::RDO$new(name = "iris_selected_columns",
               dependencies = list(iris_selected_rows))

iris_selected_columns$code <- expression({
  iris_selected_columns <- iris_selected_rows[, c("Species", "Petal.Width")]
})

iris_mtcars <-
  RDO::RDO$new(name = "iris_mtcars",
               dependencies = list(mtcars_whole,
                                   iris_selected_columns))

iris_mtcars$code <- expression({
  iris_mtcars <- cbind(iris_selected_columns, mtcars_whole)
})

iris_mtcars_test <-
  RDO::RDO$new(name = "iris_mtcars_test",
               dependencies = list(iris_mtcars))

iris_mtcars_test$code <- expression({

  stopifnot("mpg" %in% names(iris_mtcars))
  iris_mtcars <- iris_mtcars

})
```

## Deep dependencies

Most RDO methods has `deep` parameter. W
hen set to `FALSE` (default) you won't be interacting with RDO dependencies. 
If set to `TRUE` the method will touch all deep dependencies inside your RDO.
By _deep dependencies_ I mean all dependencies of the RDO, their dependencies, 
and dependencies of their dependencies, etc.

Let's take the `iris_mtcars_test` RDO that is at the top of our RDO tree,
and see the code needed to reproduce the whole tree. 

```{r}
iris_mtcars_test$print_code(deep = TRUE)
```

We can check what dependencies are validated.

```{r}
iris_mtcars_test$is_validated(deep = TRUE)
```

Let's run the code with `deep = TRUE`. 
It will run the code of all not validated RDOs.

```{r}
iris_mtcars_test$run(deep = TRUE)
iris_mtcars_test$is_validated(deep = TRUE)
```

Now the `iris_mtcars_test` RDO is deeply validated as all its dependencies are validated as well.

## Accessing dependencies

We can access any RDO dependency with `get_dependencies()` method.
With `deep = TRUE` the method returns a list of all deep unique dependencies.
Each RDO on this dependency list can be accessed directly. 

```{r}
names(iris_mtcars_test$get_dependencies(deep = TRUE))

iris_mtcars_test$get_dependencies(deep = TRUE)$data_mtcars$code
```

## Prunning cache

If we don't need to keep the data cache in all dependencies we can remove it 
using `prune_cache()` method. 
The cache won't be removed from locked RDOs.

Let's lock some RDOs and clear cache in the rest.

```{r}
as.numeric(iris_mtcars_test$get_status()$cache_size_total)
data_mtcars$lock()$is_locked()
data_iris$lock()$is_locked()
iris_mtcars_test$prune_cache(deep = TRUE)
as.numeric(iris_mtcars_test$get_status()$cache_size_total)
```

The cache from unlocked RDOs was removed and we reduced the total cache size.


## Ploting an RDO tree

We can plot the whole RDO tree. 
The blue RDOs are validated, the orange ones are not.
You can zoom in or out the interactive plot and see the details of 
each RDO. 

```{r, fig.height=6, fig.width=6}
iris_mtcars_test$plot_dependencies()
```

```{r eval=FALSE, include=FALSE}
html_name <- tempfile(fileext = ".html")
visNetwork::visSave(iris_mtcars_test$plot_dependencies(), html_name)
webshot::webshot(html_name, zoom = 2, file = "rdo_plot.png")

hexSticker::sticker(
  subplot = "rdo_plot.png", 
  package = "RDO", 
  h_fill = "white",
  h_color = "#1881C2",
  p_size = 50, 
  p_y = 1.1,
  p_color = "#1881C2", 
  s_x = .9, 
  s_y = 1.1, 
  s_width = .6,
  filename = "inst/figures/rdo_hexsticker.png")
```




